---
title: "Spatial-temporal dispersal of Ascochyta conidia"
author: "P. Melloy"
date: "13/02/2020"
output: html_document
---

# Aschochyta spore disperal

## Aims

Using the data collected from preliminary studies undertaken by Ihsan, we aim to produce a model when simulates the spread of conidia in a field within the season.  

To do this we are going to use the blackspot model provided by Art Diggle and change the dispersal parameters. We will need to revise each of the dispersal parameters to decide if they are suitable for ascochyta blight and change them according to the data we have.

Using the asco-conidia model we will simulate the spatial-temporal movement and compare it to new data obtained in future experiments. We will then use bayesian updating to improve the model before comparing the model with future trials.

```{r libraries_and_dataImport, output.lines = 20}
library(devtools)
library(cloudy)
library(readxl)
library(dplyr)
library(lme4)
library(knitr)

# update this when OneDrive decides to work
As_dat <- as.data.frame(read_xlsx("C:/Users/U8011054/USQ/SCP - Documents/Chickpea_Ascochyta/CPSporesSpatial version 2.xlsx", sheet = "analysisData"))

Astest1 <- fetch_data(url = "https://usqprd.sharepoint.com/:x:/s/SCP/EYwqdKC38rRJloCqAe5N6TMBNTTGtJAX5RaWT17XmmAHdw?e=tEbKBI")

```

```{r format_data}
# format lesion counts into integers
As_dat[,c("counts_p1","counts_p2","counts_p3","counts_p4","counts_p5", 
          "mean_count_pot", "precip","mrain", "mwd","mws")] <- 
  lapply(As_dat[,c("counts_p1","counts_p2","counts_p3","counts_p4","counts_p5", 
                   "mean_count_pot", "precip","mrain", "mwd","mws")], as.double)

# change distance to continuous variable
As_dat$dist <- as.numeric(unlist(strsplit(As_dat$distance," m")))

As_dat[,c("site","distance", "rep")] <- 
   lapply(As_dat[,c("site","distance", "rep")],as.factor)
```


From this data we hope to obtain the following.  
  - How far conidia can be dispersed in a spread event  
  - What combination of wind speed / direction / rain effects the dispersal distance.  
  
Due to the lack of reps at some of the distances we will ignore transect as a factor. We know wind direction will influence our results and we will need to accept that adds variation for which we may not be able to account for statistically.  

I will start using lmer to analyse the mean number of lesions per plant at each station. A station is defined by a pot containing five chickpea plants. The factor distance should be fit as a continuous variable, however because we don't know if the relationship between mean lesions and distance from the conidia dispersal foci is linear, we will start with distance as a categorical variable.  
Site is a categorical variable explaining the trial location. Each site may have experienced a different number of spread events, defined by the term 'rep'. Rainfall is required for conidia to disperse from the infested foci, and each rep constitutes either a simulated rainfall event, or a natural rainfall event. The following list describes the reps per site.  

```{r weatherReps}
weatherList <- list(pbc = 
        data.frame(
           rep = c(1,2,3),
           rainfall = round(c(10,10,max(As_dat$precip + As_dat$mrain, na.rm = TRUE)),4),
           WindDirection = round(unique(As_dat$mwd[As_dat$site == "pbc"])[c(1,3,4)],4),
           WindSpeed = round(unique(As_dat$mws[As_dat$site == "pbc"])[c(1,3,4)],4)
        ),
     'Horsham SPA' = 
        data.frame(
           rep = c(1,2,3),
           rainfall = round(unique(As_dat$precip[As_dat$site == "Horsham SPA"])[c(1,3,2)],4),
           WindDirection = round(unique(As_dat$mwd[As_dat$site == "Horsham SPA"])[c(1,3,2)],4),
           WindSpeed = round(unique(As_dat$mws[As_dat$site == "Horsham SPA"])[c(1,3,2)],4)
           ),
     Curyo = data.frame(
        rep = c(1,2,3),
        rainfall = round(unique(As_dat$precip[As_dat$site == "Curyo"])[c(1,2,2)],4),
        WindDirection = round(unique(As_dat$mwd[As_dat$site == "Curyo"])[c(1,2,2)],4),
        WindSpeed = round(unique(As_dat$mws[As_dat$site == "Curyo"])[c(1,2,2)],4)
     )
     
)

iterat <- 1
for(i in weatherList){
   print(kable(i, align = "c", caption = paste0("Weather during spread event at site ",names(weatherList)[iterat])))
   irerat <- iterat +1
}

```



```{r lmer_mod}

mod1 <- lmer(mean_count_pot ~ distance + (1|site/rep),
             data = As_dat)

mod2 <- lmer(mean_count_pot ~ distance + (distance|site/rep),
             data = As_dat)

anova(mod1,mod2)

summary(mod2)
coef(mod2)
```

```{r}
As_dat_pbc <- As_dat[As_dat$site == "pbc",]

mod10 <- lmer(mean_count_pot ~ distance + (1|mws:mwd),
             data = As_dat_pbc)

mod11 <- lmer(mean_count_pot ~ distance + (distance|mws:mwd),
             data = As_dat_pbc)

mod11 <- lmer(mean_count_pot ~ distance + (distance|mws:mwd:),
             data = As_dat_pbc)

anova(mod10,mod11)

summary(mod11)
mod11_co <- coef(mod11)
cbind(mod11_co$`mws:mwd`[,1],
  coef(mod11)$`mws:mwd`[,1] + coef(mod11)$`mws:mwd`[,c('distance10 m', 'distance25 m', 'distance50 m','distance75 m')])

rowSums(cbind(coef(mod11)$`mws:mwd`[,1],
coef(mod11)$`mws:mwd`[,1] + coef(mod11)$`mws:mwd`[,c('distance10 m', 'distance25 m', 'distance50 m','distance75 m')]))


2.46/5.514
3.478/7.344
5.5366/9.947

1.126/5.514
1.573/7.344
2.476/9.947

0.782/5.514
1.431/7.344
0.976/9.947

0.619/5.514
0.49/7.344
0.76/9.947


```












First I want to create a parameter for pot_rep at each specific location.

```{r}
pot_rep <- vector()
As_dat$pot_rep <- NA_real_

for(j in unique(As_dat$site)){
   dat1 <- As_dat[As_dat$site == j,]
   for(i in unique(dat1$dist_stat)[unique(dat1$dist_stat) != "NA"]){
   x1 <- seq_along(dat1[dat1$dist_stat == i,"dist_stat"])
   pot_rep <- c(pot_rep,x1)
}
As_dat[As_dat$site == j &
          As_dat$dist_stat !="NA",
       "pot_rep"] <- pot_rep
   }


As_dat[As_dat$dist_stat != "NA",]
length(x1)


```

