---
title: "Fit GAMs to dispersal patterns of _Ascochyta_ conidia"
author: "A.H. Sparks"
date: "`r Sys.Date()`"
output: github_document
---
   
## Load Libraries and Import Data

```{r libraries, message=FALSE}
library("tidyverse")
library("mgcv")
library("theme.usq")
library("ggpubr")
```

```{r import data, message=FALSE}
dat <-
   read_csv("data/CPSporesSpatial version 2.csv") %>%
   drop_na(dist_stat) %>%
   mutate(dist = as.numeric(str_replace(distance, " m", ""))) %>% 
   mutate_at(vars(site, rep, station, transect, plant_no, pot_no), factor) %>%
   group_by(site, transect, station) %>% 
   mutate(rainfall = precip + mrain)
```

## Visualise the Dispersal Data

```{r visualise_data}
ggplot(dat, aes(x = dist, y = mean_count_pot)) +
   geom_count(na.rm = TRUE, alpha = 0.5) +
   stat_summary(fun.y = "median",
                geom = "line",
                na.rm = TRUE) +
   stat_summary(
      fun.y = "median",
      colour = usq_cols("usq yellow"),
      na.rm = TRUE,
      size = 5.5,
      geom = "point",
      alpha = 0.65
   ) +
   theme_usq() +
   ylab("Median and high/low mean lesion count values") +
   xlab("Distance (m)")
```

## Fit GAMs

### mod1 - s(Distance)

```{r fit-mod1}
mod1 <-
   gam(
      mean_count_pot ~ s(dist, k = 5),
      data = dat
   )

plot(mod1)

summary(mod1)
AIC(mod1)
```

The `k` value has to be set to 5, default is 10. Else the model will not fit due
to lack of degrees of freedom.

### mod2 - s(Distance) plus Precipitation 

```{r fit-mod2}
mod2 <-
   gam(
      mean_count_pot ~ rainfall + s(dist, k = 5),
      data = dat
   )

plot(mod2)

summary(mod2)
AIC(mod2)
```

The `k` value has to be set to 5, default is 10. Else the model will not fit due
to lack of degrees of freedom.


### mod3 - s(Distance and Windspeed) plus Precipitation

```{r fit-mod3, echo=FALSE}
mod3 <-
   gam(mean_count_pot ~ rainfall + mws + s(dist, k = 5),
       data = dat)

plot(mod3)

summary(mod3)
paste("AIC: ", AIC(mod3))
```

The `k` value has to be set to 5, default is 10. Else the model will not fit due
to lack of degrees of freedom.

## Thoughts

Examining the AIC, GCV and R^2 values, the best fitting model so far is `mod2`
with the smoothed distance and precipitation. Adding windspeed decreases the
model's fitness and the graph indicates less of a good explainer for dispersal
as well.
