---
title: "Fit GAMs to dispersal patterns of _Ascochyta_ conidia"
author: "A.H. Sparks"
date: "`r Sys.Date()`"
output: github_document
---

## Load Libraries and Import Data

```{r libraries, message=FALSE}
library("tidyverse")
library("mgcv")
library("theme.usq")
library("broom")
```

```{r import data, message=FALSE}
source("R/wrangle_raw_data.R")
glimpse(dat)
```

## Visualise the Dispersal Data

```{r visualise_data}
ggplot(dat, aes(x = dist, y = mean_count_pot)) +
   geom_count(na.rm = TRUE, alpha = 0.5) +
   stat_summary(fun.y = "median",
                geom = "line",
                na.rm = TRUE) +
   stat_summary(
      fun.y = "median",
      colour = usq_cols("usq yellow"),
      na.rm = TRUE,
      size = 5.5,
      geom = "point",
      alpha = 0.65
   ) +
   theme_usq() +
   ylab("Median and high/low mean lesion count values") +
   xlab("Distance (m)")
```


## Fit GAMs

### mod1 - s(Distance)

```{r fit-mod1, echo=FALSE}
mod1 <-
   gam(
      mean_count_pot ~ s(dist, k = 5),
      data = dat
   )

summary(mod1)
```

The `k` value has to be set to 5, default is 10. Else the model will not fit due
to lack of degrees of freedom.

### mod2 - s(Distance) plus Precipitation

```{r fit-mod2, echo=FALSE}
mod2 <-
   gam(
      mean_count_pot ~ sum_rain+ s(dist, k = 5),
      data = dat
   )

summary(mod2)
```

### mod3 - s(Distance) Windspeed
```{r fit-mod3, echo=FALSE}
mod3 <-
   gam(mean_count_pot ~ mws + s(dist, k = 5),
       data = dat)

summary(mod3)
```

### mod4 - s(Distance) Windspeed plus Precipitation
```{r fit-mod4, echo=FALSE}
mod4 <-
   gam(mean_count_pot ~ sum_rain + mws + s(dist, k = 5),
       data = dat)

summary(mod4)
```

### mod5 - s(Distance + Windspeed) plus Precipitation

```{r fit-mod5, echo=FALSE}
mod5 <-
   gam(
      mean_count_pot ~ sum_rain + s(dist + mws, k = 5),
      data = dat
   )

summary(mod5)
```

## Compare the Models

### AIC, BIC
```{r compare-mods, echo=FALSE}
models <- list(mod1 = mod1, 
               mod2 = mod2, 
               mod3 = mod3,
               mod4 = mod4,
               mod5 = mod5)
map_df(models, glance, .id = "model") %>% 
   arrange(AIC)
```

### R^2^
```{r r2, echo=FALSE}
enframe(c(
   mod1 = summary(mod1)$r.sq,
   mod2 = summary(mod2)$r.sq,
   mod3 = summary(mod3)$r.sq,
   mod4 = summary(mod4)$r.sq
)) %>%
   arrange(desc(value))
```

### ANOVA
```{r anova, echo=FALSE}
anova(mod1, mod2, mod3, mod4, mod5, test = "F")
```

### Visualise the Models

```{r vis-mods, echo=FALSE}
par(mfrow = c(2, 2))
plot(mod1, main = "s(Distance)")
plot(mod2, main = "s(Distance) + Precipitation")
plot(mod3, main = "s(Distance) + Windspeed")
plot(mod4, main = "s(Distance) + Windspeed + Precipitation")
plot(mod5, main = "s(Distance + Windspeed) + Precipitation")
```

