---
title: "Spatial-temporal dispersal of _Ascochyta_ conidia"
author: "P. Melloy"
date: "`r Sys.Date()`"
output: html_document
---

# Aschochyta spore disperal

## Aims

Using the data collected from preliminary studies undertaken by Ihsan, we aim to produce a model when simulates the spread of conidia in a field within the season.  

```{r libraries_and_dataImport, output.lines = 20}
library(readxl)
library(dplyr)
library(lme4)
library(knitr)
library(mgcv)
```

```{r import_data, message=FALSE}
# import
lesion <- read_csv(here("cache", "summarised_lesion_data.csv"))
weather <- read_csv(here("cache", "weather_summary.csv"))

dat <- left_join(lesion, weather, by = c("site", "rep"))
```


From this data we hope to obtain the following.  
  - How far conidia can be dispersed in a spread event  
  - What combination of wind speed / direction / rain effects the dispersal distance.  
  
Due to the lack of reps at some of the distances we will ignore transect as a factor. We know wind direction will influence our results and we will need to accept that adds variation for which we may not be able to account for statistically.  

I will start using 'lmer' to analyse the mean number of lesions per plant at each station. A station is defined by a pot containing five chickpea plants. The factor distance should be fit as a continuous variable, however because we don't know if the relationship between mean lesions and distance from the conidia dispersal foci is linear, we will start with distance as a categorical variable.  
Site is a categorical variable explaining the trial location. Each site may have experienced a different number of spread events, defined by the term 'rep'. Rainfall is required for conidia to disperse from the infested foci, and each rep constitutes either a simulated rainfall event, or a natural rainfall event. The following list describes the reps per site.  

```{r weatherReps}
weatherList <- list(pbc = 
        data.frame(
           rep = c(1,2,3),
           rainfall = round(c(10,10,max(As_dat$precip + As_dat$mrain, na.rm = TRUE)),4),
           WindDirection = round(unique(As_dat$mwd[As_dat$site == "pbc"])[c(1,3,4)],4),
           WindSpeed = round(unique(As_dat$mws[As_dat$site == "pbc"])[c(1,3,4)],4)
        ),
     'Horsham SPA' = 
        data.frame(
           rep = c(1,2,3),
           rainfall = round(unique(As_dat$precip[As_dat$site == "Horsham SPA"])[c(1,3,2)],4),
           WindDirection = round(unique(As_dat$mwd[As_dat$site == "Horsham SPA"])[c(1,3,2)],4),
           WindSpeed = round(unique(As_dat$mws[As_dat$site == "Horsham SPA"])[c(1,3,2)],4)
           ),
     Curyo = data.frame(
        rep = c(1,2,3),
        rainfall = round(unique(As_dat$precip[As_dat$site == "Curyo"])[c(1,2,2)],4),
        WindDirection = round(unique(As_dat$mwd[As_dat$site == "Curyo"])[c(1,2,2)],4),
        WindSpeed = round(unique(As_dat$mws[As_dat$site == "Curyo"])[c(1,2,2)],4)
     )
     
)

iterat <- 1
for(i in weatherList){
   print(kable(i, align = "c", caption = paste0("Weather during spread event at site ",names(weatherList)[iterat])))
   irerat <- iterat +1
}

```

The First models I will look at are asking:  
   (i) What is the estimated mean lesions per plant as each distance from the foci, given that the conditions of each spread event is nested within each site.  
   (ii) What is the estimated mean lesions per plant as each distance from the foci, given that the conditions of each spread event is nested within each site and the distance the spore travels is dependant on the spread event at each site.  

```{r lmer_mod}

mod1 <- lmer(mean_pot_count ~ as.factor(distance) + (1|site/SpEv),
             data = dat)

mod2 <- lmer(mean_pot_count ~ as.factor(distance) + (as.factor(distance)|site/SpEv),
             data = dat)

# Compare models
anova(mod1,mod2)
```
A comparison of the two models shows us that `mod2` is much better fit given the lower AIC.  

Lets examine the results of this analysis.  
```{r mod2_summary}
summary(mod2)
```

This analysis shows there is little residual variance compared to the the variation explained by our random effects. 
Also we can see that the amount of variation explained by site is low.  
We should test that site is a significant random intercept. 
We can also note that as the distance increases so does the variance, this could be explained by the fact the spores are only going to travel further on a bearing that is the same as the wind direction.  

```{r Site_test}
# To test site we need to create a variable that describes a spread event regardless of location

mod2 <- lmer(mean_pot_count ~ as.factor(distance) + (as.factor(distance)|site/SpEv),
             data = dat)

mod3 <- lmer(mean_pot_count ~ as.factor(distance) + (as.factor(distance)|SpEv),
             data = dat)


# Compare models
anova(mod2,mod3)
```
Site is not a significant factor and we can use spread event variable instead. We will use `mod3` in the following analysis. This is shown by the lower AIC for mod3 and that the site/SpEv nesting is not significant on the chi squared test of models. We should check this is consistent for when distance is a continuous variable.  


```{r continuous_test}
# To test continous distance

mod4 <- lmer(mean_pot_count ~ distance + (distance|SpEv),
             data = dat)

# Compare models
anova(mod3,mod4)
```
`mod 3` remains a better model compared t a model where distance is defined as continuous variable this probably relates to the non-linear relationship of `distance` and `mean lesions per pot`. For completeness lets test the nesting of `site` and `SpEv` when distance is continuous.

```{r continuous_nesting_test}
# To test continous distance

mod5 <- lmer(mean_pot_count ~ distance + (distance|site/SpEv),
             data = dat)

# Compare models
anova(mod4,mod5)
```
The model where site is not included remains the better model. Therefore we can conclude that site is not a contributing factor to the distance spores can travel.  

Finally I want to check the fit of the model when distance is fit as a ordered categorical variable, as we know it is. In the shape of the regression is tested against polynomials for non-linear terms.

```{r continuous_nesting_test}
# To test continous distance

mod6 <- lmer(mean_pot_count ~ ordered(distance) + (ordered(distance)|SpEv),
             data = dat)

# Compare models
anova(mod3,mod6)
```

This last test shows `mod6` is a "better" model however the fit does not improve by much, because there is the same AIC. It also reinforces our suspicion that the relationship between number of lesions per pot and distance from the inoculumn foci is non-linear.

```{r model_summary}
summary(mod6)
```


The fixed effects give us the mean number of lesions per plant at distance zero (intercept) and this is significantly different from zero. We can also see the mean number of lesions decreases as distance from the foci increases.  

From here we should continue with a generalised additive model, which can handle non-linear terms better than a linear model with polynomials.  