---
title: "Fit Generalised Additive Models (GAMs) to the Dispersal Data"
author: "A.H. Sparks"
date: "`r Sys.Date()`"
output:
   rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fit Generalised Additive Models (GAMs) to the Dispersal Data}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width = 12, fig.height = 9)
knitr::opts_knit$set(progress = TRUE, verbose = TRUE)
```

## Import Data

```{r libraries, message=FALSE}
library("ChickpeaAscoDispersal")
library("tidyverse")
library("broom")
library("ggpubr")
library("mgcv")
library("mgcViz")

theme_set(theme_pubclean(base_size = 14))
```

## Create Data Set for GAMs

Join the `lesion_counts` data and the `summary_weather` data to create `dat` for creating GAMs.

```{r import_data, message=FALSE}
dat <- left_join(lesion_counts, summary_weather, by = c("site", "rep"))
```

```{r fonts, include=FALSE, message=FALSE, eval=FALSE}
# run only if knitting on new computer/new R installation
# !!!! Run this in a regular interactive console not this markdown file !!!!
# this is necessary to embed fonts in .eps files for EJPP
# run only if knitting on new computer
# this is necessary to embed fonts in .eps files for EJPP
library("extrafont")

if (.Platform$OS.type == "windows") {
   font_import(pattern = "arial", prompt = FALSE)
   loadfonts(device = "postscript", quiet = TRUE)
} else {
   font_import(pattern = "Arial", prompt = FALSE)
   loadfonts(device = "postscript", quiet = TRUE)
}
```

## Fit GAMs

For reproducibility purposes, use `set.seed()`.

```{r set-seed, echo=TRUE}
set.seed(27)
```

### mod1 - s(Distance)

```{r fit-mod1}
mod1 <-
   gam(
      m_lesions ~ s(distance, k = 5),
      data = dat
   )

summary(mod1)

print(p_gam(x = getViz(mod1)) +
         ggtitle("s(Distance)"),
      pages = 1)
```

### mod2 - s(Distance) + Precipitation

```{r fit-mod2}
mod2 <-
   gam(
      m_lesions ~ sum_rain+ s(distance, k = 5),
      data = dat
   )

summary(mod2)

print(p_gam(x = getViz(mod2)) +
         ggtitle("s(Distance) + Precipitation"),
      pages = 1)
```

### mod3 - s(Distance) + Wind speed
```{r fit-mod3}
mod3 <-
   gam(m_lesions ~ mws + s(distance, k = 5),
       data = dat)

summary(mod3)

print(p_gam(x = getViz(mod3)) +
         ggtitle("s(Distance) + Wind speed"),
      pages = 1)
```

### mod4 - s(Distance) + Wind speed + Precipitation
```{r fit-mod4}
mod4 <-
   gam(m_lesions ~ sum_rain + mws + s(distance, k = 5),
       data = dat)

summary(mod4)

print(p_gam(x = getViz(mod4)) +
         ggtitle("s(Distance) + Wind speed + Precipitation"),
      pages = 1)
```

### mod5 - s(Distance + Wind Speed) + Precipitation

```{r fit-mod5}
mod5 <-
   gam(
      m_lesions ~ sum_rain + s(distance + mws, k = 5),
      data = dat
   )

summary(mod5)


print(p_gam(x = getViz(mod5)) +
         ggtitle("s(Distance + Wind Speed) + Precipitation"),
      pages = 1)
```

### mod6 - s(Distance) + s(Wind Speed) + Precipitation

```{r fit-mod6}
mod6 <-
   gam(
      m_lesions ~ sum_rain + s(distance, k = 5) + s(mws, k = 5),
      data = dat
   )

summary(mod6)


print(p_gam(x = getViz(mod6)) +
         ggtitle("s(Distance) + s(Wind Speed) + Precipitation"),
      pages = 1)
```

### mod7 - s(Distance) + s(Wind Speed)

```{r fit-mod7}
mod7 <-
   gam(
      m_lesions ~ s(distance, k = 5) + s(mws, k = 5),
      data = dat
   )

summary(mod7)

print(p_gam(x = getViz(mod7)) +
         ggtitle("s(Distance) + s(Wind Speed)"),
      pages = 1)
```

### mod8 - s(Distance) + s(Wind Speed) + s(Precipitation)

```{r fit-mod8}
mod8 <-
   gam(
      m_lesions ~ s(distance, k = 5) + s(mws, k = 5) + s(sum_rain, k = 5),
      data = dat
   )

summary(mod8)


print(p_gam(x = getViz(mod8)) +
         ggtitle("s(Distance) + s(Wind Speed) + s(Precipitation)"),
      pages = 1)
```

### mod9 - s(Distance) + s(Precipitation)

```{r fit-mod9}
mod9 <-
   gam(
      m_lesions ~ s(distance, k = 5) + s(sum_rain, k = 5),
      data = dat
   )

summary(mod9)

print(p_gam(x = getViz(mod9)) +
         ggtitle("s(Distance) + s(Precipitation)"),
      pages = 1)
```

### mod10 - s(Distance) +s(Precipitation) + Wind speed

```{r fit-mod10}
mod10 <-
   gam(
      m_lesions ~ s(distance, k = 5) + s(sum_rain, k = 5) + mws,
      data = dat
   )

summary(mod10)

print(p_gam(x = getViz(mod10)) +
         ggtitle("s(Distance) + s(Precipitation) + Wind speed"),
      pages = 1)
```

### mod11 - s(Distance) + s(Wind Speed) + s(Precipitation), family = tw()

This is the same as `mod8` but using `family = tw()`, see `?family.mgcv` for more on the families. The Tweedie distribution is used where the distribution has a positive mass at zero, but is continuous unlike the Poisson distribution that requires count data.
The data visualisation shows clearly that the mean pot count data have this shape.

```{r fit-mod11}
mod11 <-
   gam(
      m_lesions ~ s(distance, k = 5) + 
         s(mws, k = 5) + 
         s(sum_rain, k = 5),
      data = dat,
      family = tw()
   )

summary(mod11)

print(p_gam(x = getViz(mod11)) +
   ggtitle("s(Distance) + s(Wind Speed) + s(Precipitation), family = tw()"),
   pages = 1)
```

### mod12 - s(Distance, bs = "ts") + s(Precipitation, bs = "ts") Wind speed, family = tw()

Try using wind speed as a linear predictor only.

```{r fit-mod12}
mod12 <-
   gam(
      m_lesions ~ s(distance, k = 5, bs = "ts") + 
         s(mws, k = 5, bs = "ts") + 
         s(sum_rain, k = 5, bs = "ts"),
      data = dat,
      family = tw()
   )

summary(mod12)

print(
   p_gam(x = getViz(mod12)) +
      ggtitle(
         "s(Distance, bs = 'ts') + s(Wind speed, bs = 'ts')\n+ s(Precipitation, bs = 'ts'), family = tw()"
      ),
   pages = 1
)
```

### mod13 - s(Distance, bs = "ts") + s(Wind speed, bs = "ts") + s(Precipitation, bs = "ts"), family = tw()

```{r fit-mod13}
mod13 <-
   gam(
      m_lesions ~ s(distance, k = 5, bs = "ts") + 
         s(mws, k = 5, bs = "ts") + 
         s(sum_rain, k = 5, bs = "ts"),
      data = dat,
      family = tw()
   )

summary(mod13)

print(
   p_gam(x = getViz(mod13)) +
      ggtitle(
         "s(Distance, bs = 'ts') + s(Wind speed, bs = 'ts')\n+ s(Precipitation, bs = 'ts'), family = tw()"
      ),
   pages = 1
)
```

This model, same structure as `mod11`, uses thin-plate splines to shrink the coefficients of the smooth to zero when possible.

## Compare the Models

### AIC, BIC
```{r compare-mods}
models <- list(mod1 = mod1,
               mod2 = mod2,
               mod3 = mod3,
               mod4 = mod4,
               mod5 = mod5,
               mod6 = mod6,
               mod7 = mod7,
               mod8 = mod8,
               mod9 = mod9,
               mod10 = mod10,
               mod11 = mod11,
               mod12 = mod12,
               mod13 = mod13
               )
map_df(models, glance, .id = "model") %>%
   arrange(AIC)
```

### R^2^
```{r r2}
enframe(c(
   mod1 = summary(mod1)$r.sq,
   mod2 = summary(mod2)$r.sq,
   mod3 = summary(mod3)$r.sq,
   mod4 = summary(mod4)$r.sq,
   mod5 = summary(mod5)$r.sq,
   mod6 = summary(mod6)$r.sq,
   mod7 = summary(mod7)$r.sq,
   mod8 = summary(mod8)$r.sq,
   mod9 = summary(mod9)$r.sq,
   mod10 = summary(mod10)$r.sq,
   mod11 = summary(mod11)$r.sq,
   mod12 = summary(mod12)$r.sq,
   mod13 = summary(mod13)$r.sq
)) %>%
   arrange(desc(value))
```

### ANOVA
```{r anova}
anova(mod1,
      mod2,
      mod3,
      mod4,
      mod5,
      mod6,
      mod7,
      mod8,
      mod9,
      mod10,
      mod11,
      mod12,
      mod13,
      test = "F")
```

### Check Best Model Fit

#### mod11 - s(Distance) + s(Wind Speed) + s(Precipitation), family = tw()

```{r mod-check11}
mod11_vis <- getViz(mod11)
check(mod11_vis,
      a.qq = list(method = "tnorm", 
                  a.cipoly = list(fill = "light blue")), 
      a.respoi = list(size = 0.5), 
      a.hist = list(bins = 10))
```


```{r save_gam_plot, include=TRUE, eval=FALSE}
# generate a new plot.gam object just for the publication (no main title)
p11 <- p_gam(x = getViz(mod11))

# save png and eps files
png(
   file = here::here("man/figures", "Fig1.png"),
   width = 640,
   height = 640,
   units = "px",
   pointsize = 14
)
print(p11, pages = 1)
dev.off()

postscript(file = here::here("man/figures", "Fig1.eps"),
           family = "Arial")
par(mar = c(5, 3, 2, 2) + 0.1)
print(p11, pages = 1)
dev.off()

embed_fonts(
   file = here::here("man/figures", "Fig1.eps"),
   outfile = here::here("man/figures", "Fig1.eps"),
   options = "-dEPSCrop"
)
```

## Thoughts

This model, _mod11_, `m_lesions ~ s(Distance) + s(WindSpeed) + s(Precipitation) - family = tw()`, is the best performing model. It cannot be used for predictions, but suitably describes the dispersal data we have on hand with the parameters used. More data would be desirable to increase the value of `k` as evidenced in the GAM checks.
