---
title: "Import Raw Lesion Count Data"
author: "A.H. Sparks"
date: "`r Sys.Date()`"
output:
   rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lesion_counts_import}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
---

```{r libraries, message=FALSE}
library("tidyverse")
library("knitr")
```

This vignette documents the raw data import, calculation of mean lesion values and exports the new data to "inst/cache" for further use in visualisation and analysis.

In the first step, the data is imported, the "m" from the `distance` column is dropped and the column is converted to numeric.
A new colum, `SpEv` (Spread Event), is created for each location and replicated event at that location. Another new column, `degrees`, is added for each transect to indicate the direction in degrees based on J. Fanning's best information.
Lastly the mean pot count, `mean_pot_count` is calculated for each of the transects at each distance, 0, 10, 25, 50 and 75 meters from the plots.

After the data are updated, a column for precipitation type, `ptype`, is added and the spread events, `SpEv`, are renamed for clarity and ordered as factors for data visualisation.

## Import and Inspect Lesion Data

```{r import_data, message=FALSE}
lesion_counts <-
   read_csv(
      system.file(
         "extdata",
         "lesion_counts.csv",
         package = "ChickpeaAscoDispersal",
         mustWork = TRUE
      )
   ) %>%
   mutate(site = str_remove(site, " SPA")) %>%
   mutate(distance = as.numeric(str_replace(distance, " m", ""))) %>%
   unite(SpEv, c(site, rep), remove = FALSE) %>%
   mutate(
      degrees = case_when(
         site == "Curyo" & transect == 1 ~ 290,
         site == "Curyo" & transect == 2 ~ 300,
         site == "Curyo" & transect == 3 ~ 310,
         site == "Curyo" & transect == 4 ~ 320,
         site == "Curyo" & transect == 5 ~ 330,
         site == "Curyo" & transect == 6 ~ 340,
         site == "Curyo" & transect == 7 ~ 350,
         site == "Curyo" & transect == 8 ~ 360,
         site == "Curyo" & transect == 9 ~ 10,
         site == "Curyo" & transect == 10 ~ 20,
         site == "Horsham" & transect == 1 ~ 45,
         site == "Horsham" & transect == 2 ~ 55,
         site == "Horsham" & transect == 3 ~ 65,
         site == "Horsham" & transect == 4 ~ 75,
         site == "Horsham" & transect == 5 ~ 85,
         site == "Horsham" & transect == 6 ~ 95,
         site == "Horsham" & transect == 7 ~ 105,
         site == "Horsham" & transect == 8 ~ 115,
         site == "Horsham" & transect == 9 ~ 125,
         site == "Horsham" & transect == 10 ~ 135,
         site == "pbc" & transect == 1 ~ 45,
         site == "pbc" & transect == 2 ~ 55,
         site == "pbc" & transect == 3 ~ 65,
         site == "pbc" & transect == 4 ~ 75,
         site == "pbc" & transect == 5 ~ 85,
         site == "pbc" & transect == 6 ~ 95,
         site == "pbc" & transect == 7 ~ 105,
         site == "pbc" & transect == 8 ~ 115,
         site == "pbc" & transect == 9 ~ 125,
         site == "pbc" & transect == 10 ~ 135
      )
   ) %>%
   mutate(
      ptype = case_when(
         SpEv == "pbc_1" ~ "irrigation",
         SpEv == "pbc_2" ~ "irrigation",
         SpEv == "pbc_3" ~ "mixed",
         SpEv == "Horsham_1" ~ "rainfall",
         SpEv == "Horsham_2" ~ "rainfall",
         SpEv == "Curyo_1" ~ "rainfall"
      )
   ) %>%
   mutate(
      site = case_when(
         site == "pbc" ~ "Horsham irrigated",
         site == "Horsham" ~ "Horsham dryland",
         site == "Curyo" ~ "Curyo"
      )
   ) %>%
   mutate(
      SpEv = case_when(
         SpEv == "pbc_1" ~ "Horsham Irrg 1",
         SpEv == "pbc_2" ~ "Horsham Irrg 2",
         SpEv == "pbc_3" ~ "Horsham Mixd 1",
         SpEv == "Horsham_1" ~ "Horsham Rain 1",
         SpEv == "Horsham_2" ~ "Horsham Rain 2",
         SpEv == "Curyo_1" ~ "Curyo Rain 1"
      )
   ) %>%
   mutate_at(vars(SpEv, site, rep), factor) %>%
   mutate(mean_pot_count = rowMeans(select(., counts_p1:counts_p5),
                                    na.rm = TRUE)) %>% 
   select(-dist_stat) %>% 
   arrange(SpEv, transect, distance)

kable(lesion_counts)
```

Save lesion count data for use in visualisation and modelling.
This only needs to be done once.

```{r save-lesion-counts, eval=FALSE}
save(lesion_counts, file = "./data/lesion_counts.rda")
```
